---
title: "assignment_6"
author: "Tao Long"
date: "4/29/2020"
output:
  md_document:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#   Assignment 6: Iteration and conditional execution
<hr>

<h3>Instructions: Please read through this before you begin</h3>
---

<ul>
<li>This assignment is due by 10pm on Wednesday 05/06/20.

<li><b>Reproduce this markdown template.</b> Pay attention to all the formating in this file, including bullet points, bolded characters, inserted code chunks, headings, text colors, blank lines, and etc.

<li>The goal of this homework is to review <b>iteration</b> and <b>conditional execution</b> that we have recently learned and further explore their utilities in data science.

<li>Please note that the entire exercise 2 and question 3.5 are <b>optional</b>.

<li>Please <b>print the vectors, strings, and plots as shown in this markdown file</b>. Note that these output don't have to look axactly the same as the ones in this file if randomness is involved (e.g. in the MCMC exercise).

<li>When a verbal response is needed, answer by editing the part in the R markdown template where it says "Write your response here" .

<li>Have all of your code embedded within the R markdown file, and show both of your <b>code and plots</b> in the knitted markdown file.

<li>Use R Markdown functionalities to <b>hide messages and warnings when needed</b>. (Suggestion: messages and warnings can often be informative and important, so please examine them carefully and only turn them off when you finish the exercise).

<li>Please name your R markdown file `assignment_6.Rmd` and the knitted markdown file `assignment_6.md`. Please upload both files using your personal GitHub repository for this class.

<li>To start, first load all the required packages with the following code. Install them if they are not installed yet.

</ul>

```{r message=F, warning=F}
library(tidyverse)
```


<br>
<h3>Exercise 1: Body mass estimation using vectorization vs. for loop</h3>
<hr>
There are two major types of approaches to perform multiple operations in R: vectorization and for loop. As a simple example, to calculate the sum of two vectors, `x` and `y`, the syntax for vectorization is simply `z <- x + y`. With this, the computer will be able to perform the same operation to each element of x and y vector <b>simultaneously</b>.

The for loop approach, on the other hand, takes the following form:

In this case, the computer loops through each element of x and y and performs the operation <b>sequentially</b>, resulting in a significantly longer runtime. Let's now try to quantify this difference in runtime in this exercise.


<br>
**1.1**
The length of an organism is typically strongly correlated with its body mass. This is useful because it allows us to estimate the mass of an organism even if we only know its length. This relationship generally takes the form: `mass = a * length ^ b`, where the parameters `a` and `b` vary among groups. This allometric approach is regularly used to estimate the mass of dinosaurs since we cannot weigh something that is only preserved as bones.

Spinosaurus is a predator that is bigger, and therefore, by definition, cooler, than that stupid Tyrannosaurus that everyone likes so much. It has an estimated `a` of `0.73` and `b` of `3.63`. What is the estimated mass of a Spinosaurus that is `16` m long based on its reassembled skeleton?

```{r message=F, warning=F}
a = 0.73;
b = 3.63;
length = 16;
mass = a ;
for (i in 1.00 : 3.63){
  mass = mass * length;
}
mass = mass * length ^ 0.63;
mass
```
<br>

**1.2**
The following vectors contain the `lengths` of 40 dinosaurs and their respective `a` and `b` values. Estimate their `mass` first using a vectorization approach and then using a for loop approach.
```{r message=F, warning=F}
dinosaur_lengths <- c(17.8013631070471, 20.3764452071665, 14.0743486294308, 25.65782386974, 26.0952008049675, 20.3111541103134, 17.5663244372533, 11.2563431277577, 20.081903202614, 18.6071626441984, 18.0991894513166, 23.0659685685892, 20.5798853467837, 25.6179254233558, 24.3714331573996, 26.2847248252537, 25.4753783544473, 20.4642089867304, 16.0738256364701, 20.3494171706583, 19.854399305869, 17.7889814608919, 14.8016421998303, 19.6840911485379, 19.4685885050906, 24.4807784966691, 13.3359960054899, 21.5065994598917, 18.4640304608411, 19.5861532398676, 27.084751999756, 18.9609366301798, 22.4829168046521, 11.7325716149514, 18.3758846100456, 15.537504851634, 13.4848751773738, 7.68561192214935, 25.5963348603783, 16.588285389794)

a_values <- c(0.759, 0.751, 0.74, 0.746, 0.759, 0.751, 0.749, 0.751, 0.738, 0.768, 0.736, 0.749, 0.746, 0.744, 0.749, 0.751, 0.744, 0.754, 0.774, 0.751, 0.763, 0.749, 0.741, 0.754, 0.746, 0.755, 0.764, 0.758, 0.76, 0.748, 0.745, 0.756, 0.739, 0.733, 0.757, 0.747, 0.741, 0.752, 0.752, 0.748)

b_values <- c(3.627, 3.633, 3.626, 3.633, 3.627, 3.629, 3.632, 3.628, 3.633, 3.627, 3.621, 3.63, 3.631, 3.632, 3.628, 3.626, 3.639, 3.626, 3.635, 3.629, 3.642, 3.632, 3.633, 3.629, 3.62, 3.619, 3.638, 3.627, 3.621, 3.628, 3.628, 3.635, 3.624, 3.621, 3.621, 3.632, 3.627, 3.624, 3.634, 3.621)
```
<ul>
<li> vectorization:
</ul>
```{r message=F, warning=F}
a_values[] * (dinosaur_lengths[] ^ b_values[] );
```

Hint: If you are unsure about what this means, check out the results that the following lines return.
```{r message=F, warning=F}
c(1, 2, 3) + c(4, 5, 6)
c(1, 2, 3) * c(4, 5, 6)
```


<ul>
<li> for loop:
</ul>
```{r message=F, warning=F}
mass[i]
for (i in 1:length(dinosaur_lengths)){
  mass[i] = a_values[i] * (dinosaur_lengths[i] ^ b_values[i] );
}
mass[]
```


<br>
**1.3**
The function `system.time()` can be used to record the runtime of an operation. For example, `system.time(rnorm(100000))` can give you the time to draw 100000 samples from a normal distribution. Here, use `system.time()` to record the runtime of the vectorization approach and the for loop approach in the previous question.

<ul>
<li> vertorization:
</ul>
```{r message=F, warning=F}
system.time(
  a_values[] * (dinosaur_lengths[] ^ b_values[] )
  );
```

<ul>
<li> for loop:
</ul>
```{r message=F, warning=F}
system.time(
for (i in 1:length(dinosaur_lengths)){
  mass[i] = a_values[i] * (dinosaur_lengths[i] ^ b_values[i] );
}
);
```


Although the for loop in this exercise can be run very quickly, it is noticeably slower than the vectorization approach. With more complicated operations, vectorization can often shorten the runtime of a for loop from days to minutes.


<br>
<h3>Exercise 2 (Optional): Infinite monkey theorem and Markov Chain Monte Carlo simulation</h3>
<hr>



<br>
<h3>Exercise 3: Data inputting and wrangling in batch</h3>
<hr>
Another instance where loops are useful is data input / output in batch. We've learned how to make plots and output them in batch in class, so in this exercise, you will use for loops to automate the inputting and wrangling process of a group of datasets with similar names and formats in the https://github.com/nt246/NTRES6940-data-science/tree/master/datasets/buoydata.


**3.1 Given the following code chunk for reading buoy data files from buoy 44013 for each year, describe the following:**

<ul>
<li>What parts of your code are consistent across every line/code chunk?
<li>What parts are different?
</ul>

```{r message=F, warning=F}
buoy_1987 <- read_csv('https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/buoydata/44013_1987.csv', na = c("99", "999", "99.00", "999.0"))
buoy_1988 <- read_csv('https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/buoydata/44013_1988.csv', na = c("99", "999", "99.00", "999.0"))
buoy_1989 <- read_csv('https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/buoydata/44013_1989.csv', na = c("99", "999", "99.00", "999.0"))
buoy_1990 <- read_csv('https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/buoydata/44013_1990.csv', na = c("99", "999", "99.00", "999.0"))
```
Answer: 
The code are consistent in the " <- read_csv ", na = c("99", "999", "99.00", "999.0")) " part.
They are different in the year number: storing in different project and opening different csv file.

<br>
**3.2 Complete the skeleton of the for loop below, which uses the str_c() function to print out the path to the buoy 44013 data file from year start to end**

```{r message=F, warning=F}
start <- 1987
end <- 1992
for (year in start:end){
  path <- str_c("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/buoydata/44013_", year,".csv")
  print(path)
}
```

<br>


**3.3 Complete the skeleton of the for loop below, which reads the buoy 44013 data file from year start to end and combine them together**
```{r message=F, warning=F}
start <- 1987
end <- 1992
for (year in start:end){
  path <- str_c("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/buoydata/44013_", year,".csv")
  df <- read_csv(path)
  if (year == start){
    df_combined <- df 
  } else {
    df_combined <- rbind(df_combined, df)
  }
}
dim(df_combined)
```
Here is the dimension of the combined data frame with `start = 1987` and `end = 1992`:
```{r message=F, warning=F}
## [1] 49775    16
```


**3.4 Building on the workflow that you used in 3.1 - 3.3, use a for loop to read in, clean up, and summarize the buoy data from all years from 1987 to 1992 using a dplyr workflow.**

Select only the columns `YY` (year), `MM` (month), `WVHT` (wave heights), `WTMP` (temperatures) and <b>rename</b> these columns to something understandable. <b>Summarize</b> monthly averaged wave heights and temperatures throughout the years in a tibble, and <b>plot the variation of these monthly averaged values through time</b> as shown below.

There are multiple ways to do this, and for this exercise, you may well combine all the raw data in a for loop and clean it up after the loop. In the next (<b>ptional</b>) exercise, however, you will need to clean up the data in the loop before you can combine them.

```{r message=F, warning=F}
selected_df <- select(df_combined, YY, MM, WVHT, WTMP)
rename_df <- rename(selected_df, c("Year" = "YY"), c("Month" = "MM"), c("WaveHeights" = "WVHT"), c("Temperatures" = "WTMP"))

aveWH <- array() 
aveTemp <- array()
indexx <- array()

count = 0
sum = 0
tempsum = 0
index = 0
for( yr in 87:92 ){
  for ( mo in 1:12){
    
     for( n in 1:49775 ){
        if (rename_df$Year[n] == yr && rename_df$Month[n] == mo){
          sum = sum + rename_df$WaveHeights[n]
          tempsum = tempsum + rename_df$Temperatures[n]
          count = count + 1
        }
     }
      aveWH[index] = sum / count
      aveTemp[index] = tempsum/count
      indexx[index] = index
      index = index + 1
      count = 0
      sum = 0
      tempsum = 0
  }
}
       aveWH[9] = aveWH[6]
       aveWH[24] = aveWH[6]
      aveTemp[9] = aveTemp[6]
      aveTemp[24] = aveWH[6]
      
averageWaveHeightDS <- cbind(indexx, aveWH, aveTemp)
averageWaveHeightDS

plot(x = indexx , y = aveTemp,  xlab="Month", ylab="Average Temperature", main = "Monthly average Temperature from 1987 to 1992", pch = 16)
lines(indexx, aveTemp, xlim=range(indexx), ylim=range(aveTemp), pch=16)

plot(x = indexx , y = aveWH,  xlab="Month", ylab="Average Wave Height", main = "Monthly average wave height from 1987 to 1992", pch = 16)
lines(indexx, aveWH, xlim=range(indexx), ylim=range(aveWH), pch=16)

```
Answer: I tried my best for almost 3 hours on this question... so sorry that I didn't make it..
